---
title: "Business Analytics and Data Mining - Week 1 Notes"
author: "Mary Anna Kivenson"
date: "February 2, 2019"
output:
  rmarkdown::html_document:
    theme: cosmo 
    highlight: zenburn
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Models with R Chapter 1  {.tabset .tabset-fade}

```{r message=FALSE, warning=FALSE}
library(faraway)
library(GGally)
library(ggplot2)
```


## Question 1.1

The dataset teengamb concerns a study of teenage gambling in Britain. Make a numerical and graphical summary of the data, commenting on any features that you find interesting. Limit the output you present to a quantity that a busy reader would find sufficient to get a basic understanding of the data. 

First, we load the data:

```{r}
data(teengamb)
```

Next, we take a look at a summary of the data:

```{r}
summary(teengamb)
```

Let's visualize the relationship of each of the features in the dataset with gamble.
It looks like:

* Income has a strong positive correlation with gamble
* Verbal has a weak negative correlation with gamble
* Status has a weak negative correlation with gamble
* Sex has a moderate negative correlation with gamble

```{r}
ggcorr(teengamb)
```


## Question 1.3

The dataset prostate is from a study on 97 men with prostate cancer who were due to receive a radical prostatectomy. Make a numerical and graphical summary of the data as in the first question.

First, we load the data:

```{r}
data(prostate)
```

Next, we take a look at a summary of the data:

```{r}
summary(prostate)
```

Let's take a look at the distributions of each of these variables:

* lcavol, lweight, age, and lpsa have somewhat normal distributions
* lbph, svi, lcp, gleason, and pgg45 have left-skewed distributions

```{r}
ggplot(prostate, aes(lcavol)) + geom_histogram(binwidth = 1)
ggplot(prostate, aes(lweight)) + geom_histogram(binwidth = 0.5)
ggplot(prostate, aes(age)) + geom_histogram(binwidth = 5)
ggplot(prostate, aes(lbph)) + geom_histogram(binwidth = 0.5)
ggplot(prostate, aes(svi)) + geom_histogram(binwidth = 0.5)
ggplot(prostate, aes(lcp)) + geom_histogram(binwidth = 0.5)
ggplot(prostate, aes(gleason)) + geom_histogram(binwidth = 0.5)
ggplot(prostate, aes(pgg45)) + geom_histogram(binwidth = 10)
ggplot(prostate, aes(lpsa)) + geom_histogram(binwidth = 1)

```


## Question 1.4

The dataset sat comes from a study entitled “Getting What You Pay For: The Debate Over  Equity  in  Public  School  Expenditures.”  Make  a  numerical  and  graphical summary of the data as in the first question.


```{r}
data(sat)
```

Next, we take a look at a summary of the data:

```{r}
summary(sat)
```

Let's visualize the relationship between expenditure, salary, and takers vs SAT score.

* The relationship between expend and total score is weak and negative. Variability does not appear to be constant for all values of expend.
* The relationship between salary and total score is weak and negative.
* The relationship between takers and total score is strong and negative, and does not apear to be linear. 

```{r}
ggplot(sat, aes(expend, total)) + geom_point()
ggplot(sat, aes(salary, total)) + geom_point()
ggplot(sat, aes(takers, total)) + geom_point()
```

## Question 1.5

The dataset divusa contains data on divorces in the United States from 1920 to 1996. Make a numerical and graphical summary of the data as in the first question.

```{r}
data(divusa)
```

Next, we take a look at a summary of the data:

```{r}
summary(divusa)
```

Let's look at time series of each of the variables. We notice that:

* Divorce spikes in 1945 and 1980
* Unemployed spikes in 1935
* Feblab spikes in 1945
* Marrige spikes in 1950
* Birth spikes in 1955
* Military spikes in 1945

```{r}
ggplot(divusa, aes(year, divorce)) + geom_step(direction = "hv") 
ggplot(divusa, aes(year, unemployed)) + geom_step(direction = "hv") 
ggplot(divusa, aes(year, femlab)) + geom_step(direction = "hv") 
ggplot(divusa, aes(year, marriage)) + geom_step(direction = "hv") 
ggplot(divusa, aes(year, birth)) + geom_step(direction = "hv") 
ggplot(divusa, aes(year, military)) + geom_step(direction = "hv") 
```

Let's look at the relationship between femlab and divorce (since when more men entered the military during WW2, more women entered the work force)

We see a spike in divorce right after the spike in women in the labor force.

```{r}
ggplot(divusa, aes(year)) + 
  geom_line(aes(y = military, colour = "military")) + 
  geom_line(aes(y = femlab, colour = "femlab"))
```

Let's look at the relationship between marriage, birth, and unemployement.

* When unemployment goes up, marriage and birth drop
* Marriage and birth are closely related except during 1950 - 1965

```{r}
ggplot(divusa, aes(year)) + 
  geom_line(aes(y = unemployed, colour = "unemployed")) + 
  geom_line(aes(y = marriage, colour = "marriage")) + 
  geom_line(aes(y = birth, colour = "birth"))
```




