---
title: "Model Moneyball"
author: "Mary Anna Kivenson"
date: "3/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(corrplot)
library(psych)
library(ggplot2)
require(gridExtra)
library(car)
library(mice)
library(VIM)
library(caret)
library(dplyr)
library(MASS)
```

#### Read Data

Here, we read the dataset and shorten the feature names for better readibility in visualizations.

```{r}
df <- read.csv("https://raw.githubusercontent.com/mkivenson/Business-Analytics-Data-Mining/master/Moneyball%20Regression/moneyball-training-data.csv")[-1]
names(df) <- sub("TEAM_", "", names(df))
names(df) <- sub("BATTING_", "bt_", names(df))
names(df) <- sub("BASERUN_", "br_", names(df))
names(df) <- sub("FIELDING_", "fd_", names(df))
names(df) <- sub("PITCHING_", "ph_", names(df))
names(df) <- sub("TARGET_", "", names(df))
head(df)
```



### Fill Missing Values

The following features have missing values.

* bt_SO - Strikeouts by batters
* br_SB - Stolen bases 
* br_CS - Caught stealing 
* bt_HBP - Batters hit by pitch (get a free base) 
* ph_SO - Strikeouts by pitchers
* fd_DP - Double Plays

Since most values in bt_HBP are missing (90%), we will drop this feature. Features br_CS and fd_DP also have many missing values, and their correlation with WINS is also weak. They will be dropped from this model as well.

```{r}
df <- subset(df, select = -c(bt_HBP, br_CS, fd_DP))
```



#### Multivariate Imputation by Chained Equations (mice)

We will use Multivariable Imputation by Chained Equations (mice) to fill the missing variables.


```{r message=FALSE, warning=FALSE, include=FALSE}
tempData <- mice(df,m=5,maxit=50,meth='pmm',seed=500)
df <- complete(tempData,1)
```


## Model Build 

### Feature Engineering

We make a few adjustments to the featurs.

* Replace bt_H (total base hits by batters) with BT_1B = bt_H - BT_2B - BT_3B - BT_HR (1B base hits)
* Apply a log transform to all of the features that do not have a normal distribution. Many of the distributions are skewed or wide, so this will help normalize the features.
* There are two features (bt_3B and ph_SO) that do not have a significant p-value when included in the linear regression model, so they will be dropped.

```{r echo=TRUE}
df2 <- df

#feature modification 
df2$bt_1B <- df$bt_H - df$bt_2B - df$bt_3B - df$bt_HR
df2 <- subset(df2, select = -c(bt_H))

#resolve skewness or wide distribution
df2$bt_3B <- log(df2$bt_3B + 1)
df2$br_SB <- log(df2$br_SB + 1)
df2$ph_SO <- log(df2$ph_SO + 1)
df2$bt_BB <- log(df2$bt_BB + 1)
df2$ph_H <- log(df2$ph_H + 1)
df2$ph_BB <- log(df2$ph_BB + 1)
df2$ph_HR <- log(df2$ph_HR + 1)
df2$fd_E <- log(df2$fd_E + 1)
df2$bt_HR <- log(df2$bt_HR + 1)
df2$bt_SO <- log(df2$bt_SO + 1)
df2$bt_2B <- log(df2$bt_2B + 1)


#drop features with low correlation
df2 <- subset(df2, select = -c(bt_3B, ph_SO))
```



### Spatial Sign

One of the significant aspects of this model is a spacial sign tranformation. Our data has very dramatic outliers, and it is difficult to detect them. Our linear regression model is highly sensitive to outliers, and a spacial sign transformaton will address that. This tranformation will project feature values onto a unit cirle, making all of the samples will be the same distance from the center of the circle This resolves the effect outliers have on our data. 

[Spacial Sign Documentation](https://rdrr.io/cran/caret/man/spatialSign.html)

In the previous step, we normalized and centered all of the features - this is a very important step that must be done before a spacial sign tranformation. We can now apply the transformation.

```{r}
df2 = as.data.frame(spatialSign(df2))
df2
```





### Multiple Linear Regression

```{r}
# Multiple Linear Regression
fit <- lm(WINS ~  ., data = df2)
summary(fit) # show results
```


